---
title: "Causal Inference Final Project: Assessment and Critique of a Study of the Causal Relationship between Air Pollution and COVID-19 Mortality"
author: "Gwen Oliver, Jiye Kwon, Hannah Gorman, Thomas Shao"
date: "12/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(questionr)
library(geepack)
library(broom)
#library(CMAverse)
library("geepack")
library("broom")
library("dplyr")
library("MASS")
library("lme4")
library("glmmTMB")
library("gamm4")
library(MatchIt)
```


```{r}
# Prior to running this whole markdown
# download the RDS file and redirect your working directory 

# read in compiled dataframe 
aggregate_pm_census_cdc_test_beds<- readRDS("./data/aggregate_pm_census_cdc_test_beds_P6.RDS") #result of pre-processing 
final_df<- aggregate_pm_census_cdc_test_beds
```

# Table 2. Controlling for confounders via propensity score analysis  

## Table 2a.Estimates based on propensity score analysis adjusting for original covariates  
```{r}
#Estimate propensity score
fitps.t2a <- glm(pm25 ~ face_masks + Tests_per_100K
             + perc_unins, data=final_df, family=binomial())
summary(fitps.t2a)
final_df$ps <- predict(fitps.t2a, final_df, type="response")

summary(final_df$ps[final_df$pm25==0])
summary(final_df$ps[final_df$pm25==1])


############################################
#2a. PS: Outcome regression
############################################
fit.t2a<-glm(data=final_df, Deaths ~ pm25 + ps
         + factor(q_popdensity)
         + scale(poverty) + scale(log(median_house_value))
         + scale(log(median_household_income)) + scale(owner_occupied) 
         + scale(no_grad) + scale(blk_pct) + scale(hispanic_pct)
         + scale(older_pecent) + scale(prime_pecent) + scale(mid_pecent) 
         + scale(date_since_social) + scale(date_since)
         + scale(beds/population) 
         + scale(obese) + scale(smoke)
         + scale(mean_summer_temp) + scale(mean_winter_temp) + scale(mean_summer_rm) + scale(mean_winter_rm)
         + offset(log(population)))

summary(fit.t2a)
cbind(coef(fit.t2a),confint(fit.t2a)) #Coefficient: 24.60 SE 5.31 

#############################################
#2a. PS: Matching 
#############################################
set.seed(2021)
m.out.t2a <- matchit(pm25 ~ face_masks + Tests_per_100K + perc_unins, data=final_df,
                     method= "nearest", ratio=1, distance= "logit", caliper=0.30, replace= F)

summary(m.out.t2a)
summary(m.out.t2a, standardize = TRUE) 


g.matches <- get_matches(m.out.t2a, data =final_df,distance = "prop.score")
dim(g.matches) #multiple rows per matched unit
head(g.matches)

#check the distribution of PS by exposure in the matched data
summary(g.matches$ps[g.matches$pm25==0]) #~=0.50 mean
summary(g.matches$ps[g.matches$pm25==1]) #~=0.52 mean

#Plot the estimated propensity score
ggplot(g.matches, aes(x = ps, fill = factor(pm25))) + geom_density(alpha = 0.2) +
  xlab('Probability of High PM2.5 Concentration') +
  ggtitle('Propensity Score Distribution by PM2.5 Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')

#Outcome regression using PS-matched data 
fitmat <- glm(Deaths ~ pm25 + ps
              + factor(q_popdensity)
              + scale(poverty) + scale(log(median_house_value))
              + scale(log(median_household_income)) + scale(owner_occupied) 
              + scale(no_grad) + scale(blk_pct) + scale(hispanic_pct)
              + scale(older_pecent) + scale(prime_pecent) + scale(mid_pecent) 
              + scale(date_since_social) + scale(date_since)
              + scale(beds/population) 
              + scale(obese) + scale(smoke)
              + scale(mean_summer_temp) + scale(mean_winter_temp) + scale(mean_summer_rm) + scale(mean_winter_rm)
              + offset(log(population)), data=g.matches)
summary(fitmat)
cbind(beta = coef(fitmat), confint(fitmat)) #pm25         28.288   16.462  40.114
```

## Table 2b:  Estimates based on propensity score analysis without adjusting for original covariates

```{r}
#Estimate propensity score
fitps.t2a <- glm(pm25 ~ face_masks + Tests_per_100K
             + perc_unins, data=final_df, family=binomial())
summary(fitps.t2a)
final_df$ps <- predict(fitps.t2a, final_df, type="response")

summary(final_df$ps[final_df$pm25==0])
summary(final_df$ps[final_df$pm25==1])


############################################
#2b. PS: Outcome regression
############################################
fit.t2b<-glm(data=final_df, Deaths ~ pm25 + ps)

summary(fit.t2b)
cbind(coef(fit.t2b),confint(fit.t2b)) #Coef: 34.51  CI (26.61, 42.40) SE 4.03

#############################################
#2b. PS: Matching 
#############################################
set.seed(2021)
m.out.t2b <- matchit(pm25 ~ face_masks + Tests_per_100K + perc_unins, data=final_df,
                     method= "nearest", ratio=1, distance= "logit", caliper=0.30, replace= F)

summary(m.out.t2b)
summary(m.out.t2b, standardize = TRUE) 


g.matches.t2b <- get_matches(m.out.t2b, data =final_df,distance = "prop.score")
dim(g.matches.t2b) #multiple rows per matched unit
head(g.matches.t2b)

#check the distribution of PS by exposure in the matched data
summary(g.matches.t2b$ps[g.matches$pm25==0]) #~=0.50 mean
summary(g.matches.t2b$ps[g.matches$pm25==1]) #~=0.52 mean

#Plot the estimated propensity score
ggplot(g.matches.t2b, aes(x = ps, fill = factor(pm25))) + geom_density(alpha = 0.2) +
  xlab('Probability of High PM2.5 Concentration') +
  ggtitle('Propensity Score Distribution by PM2.5 Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')

#Outcome regression using PS-matched data 
fitmat.t2b <- glm(Deaths ~ pm25 + ps, data=g.matches.t2b)
summary(fitmat.t2b)
cbind(beta = coef(fitmat.t2b), confint(fitmat.t2b)) #coef: 39.42   CI (30.42, 48.42) SE = 4.59
```


# Table 3. Controlling for confounders via inverse probability treatment weighting and marginal structure model    
```{r}
############################################
#pre-processing: calculating ps & weights
############################################
fitps.3a <- glm(pm25 ~ face_masks + Tests_per_100K
             + perc_unins, data=final_df, family="binomial")
summary(fitps.3a)
final_df$ps_3a <- predict(fitps.3a, final_df, type="response")


#### Create the numerator of IPTW for stablizied weight

fitpn.3a <- glm(pm25 ~ 1, data=final_df, family="binomial")
summary(fitpn.3a)
final_df$pn <- predict(fitpn.3a, data=final_df, type="response")


#### Calculate the weights for the exposed/treated and the unexposed/untreated

final_df <- final_df %>% 
  mutate(uw_t = ifelse(pm25 == 1,
                       yes = 1/ps_3a,
                       no = 1/(1-ps_3a)),
         sw_t = ifelse(pm25 == 1,
                       yes = pn/ps_3a,
                       no = (1-pn)/(1-ps_3a)))

# Mean values of unstabilized and stabilized IPW
summary(final_df$uw_t) #mean =~ 2
summary(final_df$sw_t) #mean =~ 1

############################################
#3a. Unstabilized, including only confounders
############################################
fituw.t3a <- glm(Deaths~ pm25,data=final_df,
             weights = uw_t)
summary(fituw.t3a)
cbind(beta = coef(fituw.t3a), confint(fituw.t3a)) #coef: 34.82 CI (25.93, 43.71)

############################################
#3b. Stabilized, including only confounders
############################################

fitsw.t3a <- glm(Deaths~ pm25,data=final_df,
             weights = sw_t)
summary(fitsw.t3a)
cbind(beta = coef(fitsw.t3a), confint(fitsw.t3a)) #coef: 34.82 CI (25.94, 43.70)
```


```{r}
# Estimate propensity score (INCLUDING ALL COVARIATES)
fitps.t3b <- glm(pm25 ~ face_masks + Tests_per_100K + perc_unins 
               +q_popdensity+ poverty 
               + log(medianhousevalue)
               + log(medhouseholdincome)
               +pct_owner_occ
               + education
               + pct_blk + hispanic
               + older_pecent + prime_pecent + mid_pecent
                                    + date_since_social + date_since
                                    + beds/population
                                    + obese + smoke
                                    + mean_summer_temp + 
                 mean_winter_temp+ mean_summer_rm +mean_winter_rm+ 
                 offset(log(population)),
               data = final_df,
                                             family = binomial())
summary(fitps.t3b)
final_df$ps_3b <- predict(fitps.t3b,final_df, type="response")

# Create the numerator of IPTW for stabilized weight
# model pn_e = Pr(pm25=1) in this population

fitpn.t3b <- glm(pm25 ~ 1, data=final_df,family=binomial())
summary(fitpn.t3b)
final_df$pn_3b <- predict(fitpn.t3b, final_df, type="response")



# Calculate the stabilized weights for the exposed/treated and the unexposed/untreated
final_df <- final_df %>% mutate(
  sw_e_3b = ifelse(pm25 == 1,
                yes = pn_3b/ps_3b,
                no = (1-pn_3b)/(1-ps_3b)),
  uw_e_3b = ifelse(pm25 == 1,
                yes = 1/ps_3b,
                no = 1/(1-ps_3b)))

# Mean values of unstabilized and stabilized IPW
summary(final_df$uw_e_3b) #mean =~ 27.8
summary(final_df$sw_e_3b) #mean =~ 13.9


############################################
#3c. Unstabilized, including all covariates
############################################
fituw_3b <- glm(Deaths ~ pm25,data=final_df,
             weights = uw_e_3b, family = gaussian(link = "identity"))
summary(fituw_3b)
cbind(beta = coef(fituw_3b), confint(fituw_3b))
#Total effect using unstablized weight: E            1.205  -5.08   7.499

############################################
#3d. Stabilized, including all covariates
############################################
fitsw_3b <- glm(Deaths ~ pm25,data=final_df,
             weights = sw_e_3b, family = gaussian(link = "identity"))
summary(fitsw_3b)
cbind(beta = coef(fitsw_3b), confint(fitsw_3b))
#Total effect using stablized weight: E            1.20  -5.07   7.48
```


